\documentclass[a4paper]{article}

\title{\vspace*{-30mm}HG2002:   Solutions to Tutorial 7 \\  Context and Inference}
\author{Francis Bond \url{<bond@ieee.org>}}
\date{}%2011-08-15}

\newcommand{\ans}[1]{\hfill{#1}}
%\newcommand{\ans}[1]{}

\usepackage{multicol}
%\Restriction{}
%\rightfooter{}
%\leftheader{}
%\rightheader{}
\usepackage{polyglossia}
\setmainlanguage{english}
\setmainfont[Ligatures=TeX]{TeX Gyre Pagella}
\setsansfont[Ligatures=TeX]{TeX Gyre Heros}
\usepackage{xeCJK}
\setCJKmainfont{Noto Sans CJK JP}


\usepackage{mygb4e}
\newcommand{\lex}[1]{\textbf{\textit{#1}}}
\newcommand{\eng}[1]{\textit{#1}}
\newcommand{\jpn}[1]{\textit{#1}}

\newcommand{\lx}[1]{\textbf{\textit{#1}}}
\newcommand{\ix}{\ex\it}
\newcommand{\con}[1]{\textsc{#1}}
\usepackage{url}
\usepackage[normalem]{ulem}
\newcommand{\ul}[1]{\uline{#1}}
\newcommand{\txx}[1]{\textbf{#1}}

\begin{document}
\maketitle

\begin{enumerate}

\item What are the deictic words/phrases in the following sentences?
  Identify each one as \textbf{space/time/person/social} or 
  some combination of them.

  \begin{exe}
    \ex \textit{They tried to hurt me, but he came to the rescue.}
    \\ Space: come; Time: tried, came; Person: they, me, he; 
    \ex \textit{The shop is across the street.}
    \\ Space:  across; Time: is
    \ex \textit{She was sitting over there.}
    \\ Space: over there; Time: was sitting; Person: she
    \ex \textit{It is raining out now, but I hope when you read this
      it will be sunny.}
    \\ Space: out?; Time: is raining; when you read this; Person: I, you
    \ex \textit{I was born in London and I have lived there all my
      life.}
    \\ Time: was, all my life; Person: I, my
    \ex \textit{Wilt thou not listen to me?}
    \\ Time: Wilt; Person: thou, me; Social: thou
  \end{exe}
\item Describe the features used in your own language's pronoun system, with examples.
  (e.g. for Modern English: person, number, gender)
\\ Japanese
  \begin{itemize}
  \item Referent Gender: 彼 \jpn{kare} 3p:m; 彼女 \jpn{kanojo} 3p:f
  \item Status/Solidarity: \textbf{2p}: あなた \jpn{anata} honored/formal;
  きみ \jpn{kimi} impolite/informal; おまえ \jpn{omae} very impolite/informal
  \\ \textbf{1p}: 私 \jpn{watakushi}  very formal;  私 \jpn{watashi}  formal;  ぼく
   \jpn{boku}  impolite/informal; おれ  \jpn{ore} very
   impolite/informal
   \\ watashi/anata generally preferred by women; ore/omae by men
\item Person: 私 1p; あなた \jpn{anata} 2p; 彼／彼女 3p;

\end{itemize}
\item Give two examples each of \textbf{metonymy} and
  \textbf{synecdoche}.
  \begin{itemize}
  \item \textbf{metonymy}
    \begin{itemize}
    \item \eng{\ul{The pen} is mightier than \ul{the sword}}
    \item \eng{I am tangled in \ul{red tape}}
    \end{itemize}
  \item  \textbf{synecdoche}: part for whole or vice-versa
    \begin{itemize}
    \item \eng{They had been fighting, these \ul{gray-beards} and
        half-grown boys, for three days}
    \item \eng{I was interviewed by \ul{the Straits Times}}
    \end{itemize}
  \end{itemize}
  \newpage
\item Explain what is going on here 
  \\ (from \textit{The Pink Panther Strikes Again (1976))}
  \begin{description}
    \item[Clouseau:] Does your dog bite?
    \item[Hotel Clerk:] No.
    \item[Clouseau:] [bowing down to pet the dog] Nice doggie.
      \\ {}[Dog barks and bites Clouseau in the hand]
    \item[Clouseau:] I thought you said your dog did not bite!
    \item[Hotel Clerk:] That is not my dog. 
    \end{description}
     Closeau is assuming that the dog is the Clerk's, which is the
    most salient reading --- a cooperative speaker should note if it
    is not the case; because the clerk doesn't it is anomalous (and funny)
    
%\item Make your own joke using strict/sloppy identity readings.

\item Hedges are used when you know you will flout a maxim.  Which
  maxim is flouted in the following hedges (and why)?:

\begin{exe}
% Manner:
\ex \textit{This may be a bit confusing, but I remember being in a car.}
\hfill Manner

% Quality:
\ex \textit{I may be mistaken, but I thought I saw a wedding ring on her finger.}
\hfill Quality

% \ex \textit{ I'm not sure if this is right, but I heard it was a secret ceremony in Hawaii.
%\ex \textit{ He couldn't live without her, I guess.

% Quantity:
%\ex \textit{  So, to cut a long story short, we grabbed our stuff and ran.
\ex \textit{I won't bore you with all the details, but it was an exciting trip.}
\hfill Quantity

% Relation:
\ex \textit{I don't know if this is important, but some of the files are missing.}
\hfill Relation

%Quality:
\ex \textit{As far as I know, they're married.}
\hfill Quality


% Relation:
\ex \textit{This may sound like a dumb question, but whose handwriting is this?}
\hfill Relation
% \ex \textit{  Not to change the subject, but is this related to the budget?

% Manner:
%\ex \textit{I'm not sure if this makes sense, but the car had no lights.
\ex \textit{I don't know if this is clear at all, but I think the other car was reversing.}\hfill Manner

% Quantity:
\ex \textit{As you probably know, I am afraid of dogs.}
\hfill Quality
\end{exe}

\item In the text you are
  annotating for project one try to find  examples of different kinds of reference:\\[2ex]
  \begin{tabular}{lll}
    PRONOUN: Pronominal reference & \textit{I saw a dog. \ul{It} bit me.} \\
    BRIDGE:  Bridging reference & \textit{I saw a dog.  \ul{The owner} was cute.} \\
    DEFINITE: Reference with an NP & \textit{I saw a dog. \ul{The animal} bit me.} \\
  \end{tabular} \\[2ex]
  Email the examples you found to your tutor, 
  in the following format:\\
   \begin{flushleft}
    SID: sentence \\
    ROLE: NP: target (just enough to identify it: \\
~~~~~~~~~~~~~~~~~~~~~~~~~it may come from a previous sentence)
  \end{flushleft}


\bigskip

   For example:
 
  \begin{flushleft}
    11934: If he did not take immediate action, surely the string would split in half at the middle, and he would fall.\\ 
    PRONOUN: he: Kandata \\
    DEFINITE: the string: the thinly shining spider's thread
  \end{flushleft}

\end{enumerate}

\end{document}

%%% Local Variables: 
%%% coding: utf-8
%%% mode: latex
%%% TeX-PDF-mode: t
%%% TeX-engine: xetex
%%% End: 
